---
title: "Class 12"
format: html
editor: visual
author: "Celina Shen (PID: A16673724)"
---

## Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes

Upload file to R and inspect.
```{r}
file <- read.table("rs8067378_ENSG00000172057.6.txt")

summary(file)
```

Count frequency of each genotype across samples.

```{r}
frequencies <- table(file$geno)
frequencies
```

Generate boxplot to find the median expression levels for each genotype.

```{r}
plot <- boxplot(file$exp ~ file$geno, data = file,
                 main = "Expression Levels vs Genotype", 
        xlab = "Genotype", 
        ylab = "Expression Level", 
        col = "hotpink")
plot$stats[3,]
```


A/A has a sample size of 108. A/G has a sample size of 233. G/G has a sample size of 121.
The median expression levels are A/A = 31.24847 A/A = 25.06486 G/G = 20.07363.

## Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP affect the expression of ORMDL3?

See above for boxplot.

Based on the relative expression values between A/A and G/G, the A/A genotype results in higher expression of ORMDL3 compared to G/G. The expression level of the A/G SNP demonstrates haploinsufficiency, since it results in an intermediate expression level between those of A/A and G/G. 




