---
title: "Class9 Halloween Mini Project"
format: html
author: "Celina Shen (PID: A16673724)"
---

## Running Code
```{r}

candy_file <- "candy-data.csv"
candy = read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv", row.names=1)
head(candy)
nrow(candy)
sum(candy$fruity)
```
For #1 : There are 85 candy types
For #2: There are 38 fruity types of candy


```{r}
candy["Twix", ]$winpercent
candy["Reese's Peanut Butter cup",]$winpercent
candy["Kit Kat",]$winpercent
candy["Tootsie Roll Snack Bars",]$winpercent

```
Q3: My favorite candy is Twix at 81.64291
Q4: The winpercent for Kit Kat is 76.7686
Q5: The winpercent for Tootsie rolls  is 49.6535
```{r}
#install.packages("skimr")
library(skimr)
skim(candy)
```
Q6: The variable win percent because the numbers deviate drastically from all the other Variables. 
Q7. 0 means False. 1 means True
    so in this case 0 means no chocolate, 1 means yes chocolate. 
```{r}
hist(candy$winpercent)
c = candy$winpercent[as.logical(candy$chocolate)]
f = candy$winpercent[as.logical(candy$fruity)]
 n =candy$winpercent[as.logical(candy$nougat)]
 mean(c)
 mean(f)
 mean(n)
 t.test(c,f)

```
Q8: See above for code to generate the histogram.
Q9: The histogram is not symmetrical
Q10: The center is below 50
Q11: On average, the chocolate type of candies is higher ranked than the fruit type of candies.
Q12: Yes, the p value is 2.871e -08 which is less than .05, indicating a significant difference between chocolate and fruit candies.

```{r}
head(candy[order(candy$winpercent),], n=5)
tail(candy[order(candy$winpercent),], n=5)

```
Q13: The 5 least liked candy are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters. Makes sense because these sound gross!
Q14: The 5 most liked candy are Snickers, Kit Kats, Twix, Reese's Miniatures, and Reese's Peanut Butter Cups.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()

ggplot(candy) +
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```
Q15 and 16: Code is above
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "lightpink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
Q17: NIK L Nip is ranked as the worst candy.
Q18: Reese's Peanut Butter cup is ranked as the best candy.

``` {r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)

ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
Q19. Fruity candies are ranked the highest and are cheapest.
Q20. The most expensive candies are Nik L Nip, Nestle Smarties, Ring Pop, Hershey's Krackel, and Hershey's milk chocolate. Of these, Nik L Nips are the least popular. 

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
Q22. Fruity vs chocolate, and bar vs pluribus are anti-correlated. 
Q23. Chocolate and win percent are two variables that are most positively correlated with each other. 

```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)
plot(pca$x[,1:2])
plot(pca$x[,1:2], col=my_cols, pch=16)
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```
```{r}
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```
```{r}
library(plotly)
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
Q24. The three variables in the positive direction of PC1 are fruity, hard, and pluribus. This makes sense to me because fruit candy are often assorted and hard candies.



