---
title: "Class 11"
author: "Celina Shen (PID: A16673724)"
format: html
editor: visual
---

Today, before delving into structure prediction with AlphaFold we will finish off our previous lab 10 "comparative structure analysis" section.

```{r}
library(bio3d)

id <- "1ake_A"

aa <- get.seq(id)
aa
```

```{r}
b <- blast.pdb(aa)
```

```{r}
attributes(b)
```

```{r}
head(b$hit.tbl)
```

```{r}
hits <- plot(b)
```

```{r}
attributes(hits)
```

Top hits that we like from our blast result:

```{r}
hits$pdb.id
```

```{r}
# Download related PDB files
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip= TRUE)
```

I have now found and downloaded all ADK structures in the PDB database but viewing them is difficult as they need to be aligned and superposed.

I am going to install BiocManager package from CRAN. Then I can use `BiocManager::install()` to install any bioconductor package.

```{r}
# install.packages("BiocManager")
library(BiocManager)

# BiocManager::install("msa")
```

```{r}
# Align related PDBs

pdbs <- pdbaln(files, fit = TRUE, exefile = "msa")
```

## Principal Component Analysis

```{r}
pc <- pca(pdbs)
```

```{r}
plot(pc)
plot(pc, pc.axes = c(1:2))
```

To examine in more detail what PC1 (or any PC) is capturing here, we can plot the loadings or make a wee movie (trajectory) of moving along PC1.

```{r}
mktrj(pc, pc = 1, file = "pc1.pdb")
```

## Alphafold dimer images from molstar

![](DIMER_TEST_23119_UNRELAXED_RANK_001_ALPHAFOLD2_MULTIMER_V3_MODEL_1_SEED_000.PDB-DIMER_TEST_23119_UNRELAXED_RANK_002_ALPHAFOLD2_MULTIMER_V3_MODEL_5_SEED_000.PDB-DIMER_TEST_23119_UNRELAXED_RANK_003_ALPHAFOLD2_MULTIMER_V3_MODEL_4_SEED_000.PDB-.png)

![](DIMER_TEST_23119_UNRELAXED_RANK_001_ALPHAFOLD2_MULTIMER_V3_MODEL_1_SEED_000.PDB-DIMER_TEST_23119_UNRELAXED_RANK_002_ALPHAFOLD2_MULTIMER_V3_MODEL_5_SEED_000.PDB-DIMER_TEST_23119_UNRELAXED_RANK_003_ALPHAFOLD2_MULTIMER_V3_MODEL_4_SEED_000. (1).png)




