---
title: "Class 15"
format: html
editor: visual
author: "Celina Shen (PID: A16673724)"
---

Pertussis, aka Whooping Cough, is a highly contagious lung infection caused by *B. Pertussis*.

The CDC tracks Pertussis case numbers and they can be accessed [here](https://www.cdc.gov/pertussis/php/surveillance/pertussis-cases-by-year.html?CDC_AAref_Val=https://www.cdc.gov/pertussis/surv-reporting/cases-by-year.html)

We need to "scrape" this data so we can do stuff with it in R. Let's try the **datapasta** package to do this.

```{r}
cdc <- data.frame(
  Year = c(
    1922L, 1923L, 1924L, 1925L,
    1926L, 1927L, 1928L, 1929L, 1930L, 1931L,
    1932L, 1933L, 1934L, 1935L, 1936L,
    1937L, 1938L, 1939L, 1940L, 1941L, 1942L,
    1943L, 1944L, 1945L, 1946L, 1947L,
    1948L, 1949L, 1950L, 1951L, 1952L,
    1953L, 1954L, 1955L, 1956L, 1957L, 1958L,
    1959L, 1960L, 1961L, 1962L, 1963L,
    1964L, 1965L, 1966L, 1967L, 1968L, 1969L,
    1970L, 1971L, 1972L, 1973L, 1974L,
    1975L, 1976L, 1977L, 1978L, 1979L, 1980L,
    1981L, 1982L, 1983L, 1984L, 1985L,
    1986L, 1987L, 1988L, 1989L, 1990L,
    1991L, 1992L, 1993L, 1994L, 1995L, 1996L,
    1997L, 1998L, 1999L, 2000L, 2001L,
    2002L, 2003L, 2004L, 2005L, 2006L, 2007L,
    2008L, 2009L, 2010L, 2011L, 2012L,
    2013L, 2014L, 2015L, 2016L, 2017L, 2018L,
    2019L, 2020L, 2021L, 2022L, 2024L
  ),
  No..Reported.Pertussis.Cases = c(
    107473, 164191, 165418, 152003,
    202210, 181411, 161799, 197371,
    166914, 172559, 215343, 179135, 265269,
    180518, 147237, 214652, 227319, 103188,
    183866, 222202, 191383, 191890, 109873,
    133792, 109860, 156517, 74715, 69479,
    120718, 68687, 45030, 37129, 60886,
    62786, 31732, 28295, 32148, 40005,
    14809, 11468, 17749, 17135, 13005, 6799,
    7717, 9718, 4810, 3285, 4249, 3036,
    3287, 1759, 2402, 1738, 1010, 2177, 2063,
    1623, 1730, 1248, 1895, 2463, 2276,
    3589, 4195, 2823, 3450, 4157, 4570,
    2719, 4083, 6586, 4617, 5137, 7796, 6564,
    7405, 7298, 7867, 7580, 9771, 11647,
    25827, 25616, 15632, 10454, 13278,
    16858, 27550, 18719, 48277, 28639, 32971,
    20762, 17972, 18975, 15609, 18617,
    6124, 2116, 3044, 23544
  )
)
head(cdc)
```

```{r}
library(ggplot2)

baseplot <- ggplot(cdc) +
  aes(Year, No..Reported.Pertussis.Cases) +
  geom_point() +
  geom_line()

baseplot
       
```

Let's look at cases before and after the wP vaccine was rolled out (1946) and when we switched to the aP vaccine (1996). Also labeling 2020 for the significant public health events and a major outbreak during 2004.

Pertussis cases rose after the switch to the aP vaccine. This may be because the aP vaccine loses efficacy over time compared to the wP vaccine.

```{r}
baseplot +
  theme_bw() +
  geom_vline(xintercept = 1946, color = "blue") +
  geom_vline(xintercept = 1996, color = "red") +
  geom_vline(xintercept = 2020, color = "gray") +
  geom_vline(xintercept = 2004, color = "green")
```

## CMI-PB (Computational Models of Immunity - Pertussis Boost)

This project collected and makes freely available data about the immune response to Pertussis vaccination.

You can access the data via an API which returns JSON format (key:vale pairs).

We can use the **jsonlite** package and it's `read_json` function.

```{r}
library(jsonlite)

subject <- read_json("http://cmi-pb.org/api/v5/subject",
                     simplifyVector = TRUE)
head(subject)
```

> Q. How many subjects are in this dataset?

```{r}
nrow(subject)
```

> Q. How many males/females are in this dataset?

```{r}
table(subject$biological_sex)
```

> Q. How many wP and aP are in this dataset?

```{r}
table(subject$infancy_vac)
```

> Q. Break down of biological sex and race?

```{r}
table(subject$race, subject$biological_sex, subject$ethnicity)
```

> Q. Does this break down reflect the US population?

```{r}
table(subject$dataset)
```

```{r}
specimen <- read_json("http://cmi-pb.org/api/v5/specimen", simplifyVector = TRUE)
ab_titer <- read_json("http://cmi-pb.org/api/v5/plasma_ab_titer", simplifyVector = TRUE)

```

```{r}
head(specimen)
```

```{r}
head(ab_titer)
```

We want to merge or "join" these tables so we can have all the info we need about a given antibody measurement.

```{r}
#/ message
library(dplyr)

meta <- inner_join(subject, specimen)
```

```{r}
head(meta)
```

and one last join of `ab_titer` and `meta`

```{r}
abdata <- inner_join(ab_titer, meta)
```

```{r}
head(abdata)
```

```{r}
nrow(abdata)
```

```{r}
table(abdata$isotype)
```

```{r}
table(abdata$antigen)
```

Filter `abdata` for isotype IgG.

```{r}
igg <- filter(abdata, isotype == "IgG")
head(igg)
```

Make a boxplot of IgG antigen levels - this will be a plot of normalized MFI vs antigen.

```{r}
ggplot(igg) +
  aes(x = log(MFI_normalised), y = antigen, col = infancy_vac) +
  geom_boxplot()
```

Ideally, I would like to see how these Ab levels change over time relative to the booster shot.

```{r}
table(abdata$visit)
```

```{r}
ggplot(igg) +
  aes(x = log(MFI_normalised), y = antigen, col = infancy_vac) +
  geom_boxplot() +
  facet_wrap(~visit)
```

```{r}
igg_21 <- filter(igg, dataset == "2021_dataset")
table(igg_21$dataset)
```

```{r}
filter(igg, antigen == "PT", dataset == "2021_dataset")%>%
  ggplot()+
  aes(x = planned_day_relative_to_boost,
      y = MFI_normalised,
      col = infancy_vac)+
  geom_point()+
  geom_line()+
  geom_vline(xintercept = 0, linetype = "dashed")+
  geom_vline(xintercept = 14, linetype = "dashed")+
  labs ( x= "Day relative to boost", y = "Normalized MFI")
```



```{r}
abdata.21 <- abdata %>% filter(dataset == "2021_dataset")

abdata.21 %>% 
  filter(isotype == "IgG",  antigen == "PT") %>%
  ggplot() +
    aes(x=planned_day_relative_to_boost,
        y=MFI_normalised,
        col=infancy_vac,
        group=subject_id) +
    geom_point() +
    geom_line() +
    geom_vline(xintercept=0, linetype="dashed") +
    geom_vline(xintercept=14, linetype="dashed") +
  labs(title="2021 dataset IgG PT",
       subtitle = "Dashed lines indicate day 0 (pre-boost) and 14 (apparent peak levels)")
```




