---
title: "BGGN 123 Shen lab 6"
format: html
author: "Celina Shen (PID: A16673724)"
---

# Intro to functions lecture activities

```{r}
add <- function(x,y) {
  x+y
}
```

Can I just execute this chunk? No, because the "add" chunk has not been run yet!

```{r}
add (1,1)
add (c(100,1),1)
```

```{r}
add <- function(x, y=1) {
  x+y
}
```

If y≠1 in the input, the y=1 line will be overrode.

```{r}
add(5)
add (1,3)
add (c(100,1),2)
```

```{r}
add <- function(x, y, z) {
  x+y+z
}
add(10, 1, 1)
```

Make a function "generate_DNA()" that makes a random nucleotide sequence of any length.

```{r}
#generate_DNA <- function() {
bases <- c("A", "C", "G", "T")
sample(bases, size = 5, replace = TRUE)
```

Now that this part is working, I can make it into a function.

```{r}
generate_DNA <- function(length) {
  bases <- c("A", "C", "G", "T")
  sequence <- sample(bases, size = length, replace = TRUE)
  return(sequence)
}
generate_DNA(10)
```

```{r}
#install.packages("bio3d")
library(bio3d)
unique(bio3d::aa.table$aa1)[1:20]
```

Generate random protein sequences of length 6 to 12.

```{r}
generate_protein <- function(length){
  amino_acids <- unique(bio3d::aa.table$aa1)
  sequence <- sample(amino_acids, size=length, replace = TRUE)
  sequence <- paste(sequence, collapse = "")
  return(sequence)
}
```

Sequences from length 6 to 12.

```{r}
answer <- sapply(6:12, generate_protein)
answer
```

Run function

```{r}
generate_protein(6)
```

```{r}
paste(c("barry", "alice", "amy", "chandra"),
      "loves R")
```

```{r}
paste(">id.", 6:12, "\n", answer, "\n", sep ="")
```

```{r}
cat(paste(">id.", 6:12, "\n", answer, "\n", sep =""), sep = "")
```




